---
title: "Displacement of fishing effort by Large Scale Marine Protected Areas[^*]"
subtitle: "Updated on `r Sys.Date()`"
author: |
  Juan Carlos Villaseñor-Derbez^1^
  John Lynham^2^
abstract: "Large-scale Marine Protected Areas (LSMPAs) have seen a significant increase over the last years. Fishing effort is effectively eliminated within these protected areas upon implementation. The benefits of reducing effort have been largely studied, and include increases in abundance, biomass, and diversity within the bounded regions. These no-take zones may produce spillover effects, which provide fish for outside areas. However, the economic and ecological implications of displacing fishing effort are not yet fully understood. Novel data products that track fishing effort at the vessel-level allow us to identify changes in fleet- and vessel-level behavior upon the implementation of protected areas, as well as how these redistribute. This papers evaluates the implications of implementing LSMPA, by evaluating changes in fishing hours, showing that vessels in the effected region reduce fishing effort after the implementation of PIPA. Our results are robust to a set of specifications. We also track the relative spatial allocation of fishing events thorugh time, and identify that areas closer to PIPA show an increase in relative fishing hourse due to the displacement of PIPA-fishing vessels. Our results not only provide an impact evaluation of the effect of LSMPAs on fishing activity, but provide insights into vessel redistribution dynamics, which may have ecological and economic implications."
date: "
^1^Bren School of Environmental Science and Management, University of California Santa Barbara, Santa Barbara, CA\\newline
^2^Department of Economics, University of Hawaii at Manoa, Honolulu, HI"
output:
  bookdown::pdf_document2:
    fig_height: 3.5
    fig_width: 6
    toc: no
    citation_package: natbib
    keep_tex: yes
    number_sections: true
fontsize: 11pt
header-includes:
  - \usepackage{float}
  - \floatplacement{figure}{H}
  - \usepackage{lineno}
  - \linenumbers
  - \usepackage[bottom]{footmisc}
  - \usepackage{pdflscape}
  - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
bibliography: references.bib
biblio-style: chicago
---

[^*]: Work in progress, do not circulate

\clearpage

```{r, echo = F, message = F, warning = F}
suppressPackageStartupMessages({
  library(startR)
  library(sf)
  library(tidyverse)
  library(magrittr)
})

knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      eval = T)

options(knitr.table.format = "latex")

source(here::here("scripts", "st_rotate.R"))
source(here::here("scripts", "sfc_as_cols.R"))
```


# Introduction

Marine Protected Areas (MPAs) are intended to safeguard parts of the ocean from fishing and other extractive activities. Current international goals aim to protect 10\% of the ocean environments by 2020. In an effort to meet this target, the world has seen a rapid increase in MPA coverage \citep{wood_2008,sala_2018}, largely driven by a small number of Large Scale Marine Protected Areas (LSMPA) \citet{singleton_2014,boonzaier_2016,alger_2017}). Due to weak property rights, limited habitat transformation, and potentially lower management costs, LSMPAs provide an opportunity to safeguard the oceans \citet{game_2009}. Today, a small number of LSMPAs make up at least 80% of the managed areas in the ocean \citet{toonen_2013}. LSMPAs may be instrumental in helping us achieve the 10\% target by 2020 \citep{mccauley_2016}.

Given the relatively recent establishment of most LSMPAs, very little is known about their human dimensions and implication for fisheries \citep{gray_2017}. As with customary MPAs, it is important that we understand the socioeconomic implications of management interventions. One issue of particular importance is that of the displacement or redistribution of fishing effort, which may influence the outcomes of an MPA \citep{smith_2003}. Theoretical models make different assumptions about the way in which fishers will reallocate fishing effort after an area closure. The few empirical works focus on small-scale fisheries and rarely incorporate spatially explicit data. Despite these efforts, marine conservation is yet to identify the response in fishing effort and its management implications.

Recent advances in satellite tracking technologies and near real-time identification of fishing activity provide us with an opportunity to tackle increase governance and answer long-lasting questions. Here, we ask, how do fishers respond to the implementation of LSMPA? After a closure, and if they continue to fish, where do they reallocate fishing effort? We use identification of fishing activity via Automatic Identification Systems (AIS) and causal inference techniques to describe the behavioral changes and spatial redistribution of the industrial tuna purse seine fleet due to the implementation of a Large Scale Marine Protected Areas in the Pacific Ocean.

Our work is novel in the sense that it provides empirical evidence of the effect of Large Scale Marine Protected Areas in fishing behavior and distribution and can help guide future interventions as countries march towards meeting the targeted 10\% of ocean protection. Understanding how effort is displaced from LSMPAs might provide insights into how the industrial fleet would react to a high seas closure.

The paper is outlined as follows: Section \ref{background} provides more information on Large Scale Marine Protected Areas and background on empirical studies of effort redistribution. An overview of the Nauru Agreement and associated countries, a description of the fleet that operates in the region, and a brief history of PIPA is also included. Section \ref{methods} describes our data and identification strategy. Section \ref{results} presents our results, section \ref{discussion} provides an extension of our results to other LSMPAs\footnote{Not yet. but I think this would be interesting} and discusses our results\footnote{I might also include a section, after the background, where I explain the different theoretical models of effort redistribution and behavioral responses}.

# Background

## Large Scale Marine Protected Areas

The cutoff at which an MPA is considered to be a LSMPA ranges from areas larger than 30,000 km^2^ as defined by \citet{desanto_2013} or areas larger than 250,000 km^2^, as defined by \citep{toonen_2013}. Figure \ref{fig:LSMPAs_map} shows LSMPAs that meet the later, and that are fully no take. LSMPAs are often implemented in the pelagic environment, where the dominant human activity is industrial fishing \citep{gray_2017,kroodsma_2018}. The early literature on LSMPAs focused on the inherent challenges and difficulties that come with a pelagic environment. \citet{kaplan_2010} claimed that very large MPAs would result in excessive opportunity costs and that these would be difficult to enforce. \citet{game_2009} suggested that most of the challenges could be overcome with the incorporation of technology, in what then became known as Dynamic Ocean Management \citep{maxwell_2015}. The effectiveness and difficulties of implementating LSMPAs are not the focus of this paper \footnote{But see \citet{singleton_2014}, who provide an objective discussion of pros and cons of LSMPAs.}. Instead, we focus on the effect and implications of the ones that *have been* implemented.

```{r, fig.cap = "\\label{fig:LSMPAs_map}Large Scale Marine Protected Areas. The map shows all areas larger than 250,000 Km^2^. Areas in dark blue meet at least one of these conditions: reported no take area is greater than 0, are categorized as IUCN Ia or Ib, their designated english name is 'Protected Area'."}

mpas <- read_sf(dsn = here::here("data", "spatial", "LSMPAs"), layer = "LSMPAs") %>% 
  st_rotate() %>% 
  filter(!desig_type == "International") %>% 
  mutate(strict = ifelse(no_tk_area > 0 |
                           iucn_cat %in% c("Ia", "Ib") |
                           desig_eng == "Protected Area",
                         "No-Take", "Others"),
         strict = ifelse(wdpaid == 309888, "PIPA", strict),
         Legend = fct_relevel(strict, c("No-Take", "PIPA", "Others"))) %>% 
  filter(!name == "Longline")

coast <- map_data('world', wrap=c(-20,340), ylim=c(-85,75))%>%
  select(group, lat, long) %>%
  st_as_sf(coords = c("long", "lat"), crs = 4326) %>%
  group_by(group) %>% 
  summarize(geometry = st_combine(geometry)) %>%
  st_cast("POLYGON") %>% 
  mutate(a = 1) %>% 
  group_by(a) %>% 
  summarize() %>% 
  ungroup()

pipa <- sf::read_sf(dsn = here::here("data", "spatial", "PIPA"), layer = "PIPA") %>% 
  sf::st_transform(crs = "+proj=longlat +datum=WGS84 +no_defs") %>% 
  st_rotate()

ggplot() +
  geom_sf(data = coast, fill = "#E3E3E3") +
  geom_sf(data = mpas, fill = "transparent", color = "black", size = 1) +
  geom_sf(data = mpas, aes(fill = Legend), color = "transparent") +
  ggtheme_map() +
  scale_fill_manual(values = c("steelblue", "red", "steelblue1")) +
  theme(legend.justification = c(0.5, 1),
        legend.position = c(0.5, 0.05),
        legend.box = "horizontal",
        legend.background = element_blank()) +
  guides(fill = guide_legend(ncol = 3))
```

LSMPAs were erroneously assumed to have little social implications due to their remoteness. However, there have been calls to incorporate the human dimensions into LSMPAs management and evaluation \citep{agardy_2011,gray_2017}. Most research incorporating these dimensions has focused on governance and enforcement of LSMPAs (*i.e.* \citet{alger_2017,christie_2017}), but they are yet to be the focus of economic analyses \citep{gray_2017}. Overall, there has been little empirical work regarding LSMPAs. Recent technological advances in vessel-detection systems allows for the discovery and advancement of many important facets of LSMPAs. For example, \citep{mcdermott_2018} show that the anticipation of a LSMPA can lead to preemptive overfishing, which can erode or delay the expected benefits of the intervention. \citet{white_2017} combine shark tags and vessel-tracking data to demonstrate that the fairly large Palmyra Atoll National Wildlife Refuge (54,000 Km^2^) protectes two thrids of the tagged grey reef sharks by effectively excluding fishing effort. More recently, \citep{bradley_2018} use similar data to highlight cases of potential illegal shark fishing *inside* a 2 million km^2^ shark sanctuary. To date, no studies have evaluated the displacement of fishing effort due to LSMPA implementation.

Spatial closures of this magnitude are likely to induce changes in fishers' behavior. Theoretical models of fishing effort redistribution range from the simplistic assumption that effort inside the bounded region disappears, to spatially explicit models that reallocate fishing effort based on habitat characteristics, presence of other vessels, and expected returns \citep{smith_2003,hilborn_2006}. However, these focus on the long term optimal equilibrium, and redistribution of fishing effort may not always be optimal, especially over the first years \citep{stevenson_2013}.

The empirical research that has been done in custumary sized MPAs suggest that resource users may show idiosyncratic responses. For example, \citet{stevenson_2013} show that a network of MPAs displaced fishing effort farther away from ports, resulting in higher *perceived* costs, and increases in catch per unit effort. \citet{cabral_2017} analyse the redistribution of fishing and non-fishing vessels following the implementation of a network of MPAs in California, and find that commercial dive boats follow a fishing-the-line pattern, while some fishing boats follow an ideal free distribution. More recently \citet{elahi_2018} used satellite tracking data to show that a temporal spatial closure caused trawlers to maintain effort but apply it more intensively elsewhere, particularly along the borders and closer to shore. The way in which fishers react to a spatial closure can have major implications in its outcome \citep{smith_2003,hilborn_2006} highlighting the need to understand how fishers react to the implementation of LSMPAs, and fishing effort changes and is spatially redistributed. However, all these closures took place within Exlcusive Economic Zones, where other regulations exist. This may not always be the case for LSMPAs, where often the entire EEZ is converted into a LSMPA, leaving fishers with the options of moving to the high seas or other countries' EEZs.

## Nauru agreement and the Phoenix Island Protected Area

The Nauru Agreement was established in 1982 by a select group of Pacific island nations to manage their important tuna resources. PNA Members include Federated States of Micronesia, Kiribati, Marshall Islands, Nauru, Palau, Papua New Guinea, Solomon Islands, and Tuvalu. The Nauru Agreement regulated access of foreign vessels (*i.e.* those from non-PNA countries). Holding ~80% of the historical purse seining grounds within their Exclusive Economic Zones, PNA countries gained bargaining power when providing access to foreign fleets \citep{havice_2010}.

The cooperation that emerged thanks to the PNA allowed for subsequent agreements that strengthened fisheries management, like the Palau Agreement, which limited the number of purse seiners at 205 vessels from 1995-2007[^a]. However, the most notable regulation is their approach to manage fishing effort: a Vessel Day Scheme (VDS) implemented in 2007 \citep{havice_2013}. This effectively modified how fishing effort was managed, from number of vessels under the Palau Agreement to fishing hours. The VDS works as follows: Each year, scientific advisors recomend a total number of fishing vessel-days per year. Hours are allocated to each PNA country based on catch history, and they then use or sell fishing rights to other non-PNA countries \citep{aqorau_2018}. While the effectiveness of this scheme has been debated in terms of meeting their fishery management and conservation objectives, the licensing significantly contributes to the economy of these island nations \citep{havice_2010}.

[^a]: See \citet{havice_2010} for a detailed description of the Nauru, Palau, and Federal States of Micronesia agreements, their objectives, and outcomes.

```{r, fig.cap = "\\label{fig:PNA_map}Map of the Exclusive Economic Zones (EEZs) of the region of interest. A solid line indicates countries that belong to the PNA, while a dashed line indicates all others. A red line indicates the Kiribati EEZ, and a solid red polygon delineates PIPA."}
countries <- c("PIPA", "KIR", "HS", "ASM", "COK", "FSM", "MHL", "NRU", "PNG", "SLB", "TKL", "TUV", "UMI")

small_coast <- rnaturalearth::ne_countries(scale = "large", returnclass = "sf") %>% 
  filter(sov_a3 %in% countries) %>% 
  st_rotate()

eez <- read_sf(dsn = here::here("data", "spatial", "EEZ_subset"),
               layer = "EEZ_subset") %>% 
  filter(ISO_Ter1 %in% countries) %>% 
  mutate(KIR = ISO_Ter1 == "KIR") %>% 
  group_by(ISO_Ter1, PNA, KIR) %>% 
  summarize()

ggplot() +
  geom_sf(data = eez, aes(color = KIR, linetype = !PNA), fill = "transparent") +
  geom_sf(data = pipa, fill = "red") +
  geom_sf(data = small_coast, size = 1, color = "black") +
  ggtheme_plot() +
  scale_color_manual(values = c("black", "red")) +
  theme(legend.position = "none")
```

The main tuna species caught in the region are skipjack (*Katsuwonus pelamis*), yellowfin (*Thunnus albacares*), albacore (*Thunnus alalunga*) and bigeye (*Thunnus obesus*). From these, the first two are amongst the top-10 species represented in global fisheries production statistics, with 2016 catches increasing relative to the 2005-2014 average \citep{fao_2018}. Theis region of the Pacific has historically accounted for a large portion of tuna catches \citep{aqorau_1997}. Today, the PNA controls close to 50% of the global skipjack tuna production \citep{pna_website_2018}. A large portion of these catches derive from purse seine vessels licensed under the VDS. Fishing vessels from Australia, New Zeland, China, France, Korea, Japan, the Philippines, Taiwan, and the United States participate in the purse-seining VDS.

One of the most notable and recent management interventions in the region is the implementation of the Phoenix Island Protected Area (PIPA) by the government of Kiribati. PIPA was first declared in 2006, and established in 2008 with only 4% of it was declared as no-take. In January 1^st^, 2015, the no-take area within PIPA was expanded to a total area of 397,447 km^2^, roughly 1.5 times the size of Ecuador. Figure \ref{fig:PNA_map} shows a map of the PNA countries and the Phoenix Island Protected Area.

The closure of such a large area in one of the most important fishing regions in the world provides a great opportunity to evaluate the behavioral responses and redistribution of fishing effort by vessels that used to fish there. PIPA has been the focus of previous research, showing that fishing effort is effectively reduced after implementation \citep{mccauley_2016,mcdermott_2018}. To this, we pose two questions: How do individual vessels respond to the sudden exclusion of such a big area? And where did all the vessels go? In the next sections we describe the data and methods used to answer this questions.

# Methods

This section is divided into two main parts. First, we provide a general description of AIS data and the process of identification of vessel-level fishing events done by Global Fishing Watch[^1]. Alongside, we describe the subset of data used on our analyses. When relevant, we also point out possible shortcomings in the data, or factors that must be considered in the later analyses. We then move on to explain our empirical strategy for the identification of the behavioral changes and redistribution of fishing effort.

[^1]: Global Fishing Watch: [globalfishingwatch.org](globalfishingwatch.org)

## Data

Automatic Identification Systems (AIS) are on-board devices that provide at-sea safety and prevent ship collisions by broadcasting vessel position, course, and activity to surrounding vessels. These broadcasted messages can be received by satellites and land-based antennas. GFW uses convolutional neural networks to infer vessel characteristics and whether each broadcasted position represents a fishing event, thus allowing us to estimate near real-time fishing events globally since 2012 \citep{kroodsma_2018}.

The amount of data gathered is dependent on the number of antennas and satellites that can pick up the signals. In June 1^st^ 2014 satellite count increased from 3 to 6, and then to 10 in January 1^st^ 2016. This causes an increase in the number of *received* AIS messages (*i.e.* points), and therefore an apparent increase in the number of vessels and fishing hours. However, the addition of satellites "affects" all vessels in the same way. The variability in AIS data and ocean conditions require that temporal trends be taken into account. We do that by including specific controls in our identification strategy and using a subset of data that meet a BACI design -which gives us the full tracks for vessels affected and unaffected by the implementation of PIPA.

```{r, eval = T}
effort_by_vessel <- readRDS(file = here::here("data", "effort_by_vessel.rds")) %>% 
  filter(year < 2018,
         gear == "purse_seines")

effort_by_vessel_JPN <- readRDS(file = here::here("raw_data", "JPN_effort_by_vessel.rds"))
```

```{r, eval = T}
# Just to get pipa vessels that fished inside pipa before 2015
pipa_vessels <- effort_by_vessel %>% 
  filter(treated,
         baci_strict) %$%
  unique(mmsi) %>% 
  length()
```

Our analysis focuses on purse seine vessels, the most important fishery for PNA countries\footnote{We perform some of the same analyses for longliners and include them in the Appendix}. We identify a total of 103 purse seiners that fished in PNA waters at least once before 2015\footnote{New vessels that entered the fishery after 2015 were not exposed to the policy intervention in the pre-treatment period and are therefore exlcuded from our analyses.}. These vessels represent over 26 million individual observations for the 2012 - 2017 period. We identify 65 vessels that have fished inside PIPA at least once since 2012. From these, 62 did so before the anouncement (*i.e.* 09/01/2014 *sensu* \citep{mcdermott_2018}) but are only able to track 61 for the complete periof of study before and after its implementation\footnote{1 vessel never fished again after August 18, 2013.}. On the other hand, 38 vessels never fished inside PIPA but we only have data for 26 vessels before and after.

Therefore, our treatment group contains all purse seiners (n = 61) that fished within PIPA at least once before the anouncement, and that continued to fish elsewhere after the January 2015 implementation. Vessels in the control group meet the following two conditions: i) vessels never fished within PIPA waters, and ii) vessels have fished in surrounding areas (*i.e.* PNA-countries' EEZ) before and after PIPA closure (n = 26).

We include three additional definitions of control groups as a robustness check: one with only vessels that belong to PNA countries (n = 7), and one that excludes Chinese vessels (n = 21). Our third control is made up of Japanese purse seiners that fish in the Pacific but that never fished inside PIPA (n = 27)\footnote{Using Taiwanese vessels was the original idea, but there are only 4 vessels with pre-2015 data. I need to further identify propper Taiwanese vessels, as I am currently using the entire Pacific ocean}. Our definition of treatment and control groups leaves us with 61 and 21 treated and control vessels, which have just over 22 million observations where about 22% are identified as fishing events.

For each vessel we calculate total daily fishing hours and obtain pannel data with 37,800 observations. Table \ref{tab:baci_n_s} shows the number of vessels following a BACI design, as well as the fishing hours, before and after PIPA. Fig. \ref{fig:all_vessels} shows that mean fishing hours for purse seiners have an abrupt increase, just before January 1st, 2015. This trend is observed for both treated and control groups. Across all measures, the treatment and control vessels follow similar patterns, confirming our claim that the control group provides a plausible counterfactual. Figure \ref{fig:baci_strict} provides a visual representation of the vessel-level fishing events that make up each group through time. We use this data to answer our key questions: How do the 61 purse seiners modify their behavior as compared to the different control groups, and where do they go after the spatial closure? The following section describes our empirical identification strategies.

```{r, eval = T}
hours <- effort_by_vessel %>%
  filter(baci_strict) %>%
  group_by(treated, post, date, mmsi) %>% 
  summarize(hours = sum(hours, na.rm  = T)) %>% 
  ungroup() %>% 
  group_by(treated, post) %>% 
  summarize(hours = mean(hours, na.rm = T)) %>% 
  spread(post, hours) %>% 
  mutate(change = `TRUE` / `FALSE`)

# I need to group the B - A columns into something that says "hors"

effort_by_vessel %>% 
  filter(baci_strict) %>%
  group_by(treated, mmsi) %>% 
  count() %>% 
  ungroup() %>% 
  group_by(treated) %>% 
  summarize(boats = n()) %>% 
  ungroup() %>% 
  left_join(hours, by = c( "treated")) %>% 
  mutate(treated = ifelse(treated, "Treatment", "Control")) %>% 
  knitr::kable(col.names = c("Group", "n" ,"Before", "After", "Change (A / B)"),
               caption = "\\label{tab:baci_n_s}Number of fishing vessels and mean daily fishing hours by group before and after PIPA implementation.",
               booktabs = T,
               digits = 2) %>% 
  kableExtra::kable_styling(latex_options = "HOLD_position")
```

```{r, echo = F, fig.cap = "\\label{fig:all_vessels}Fishing hours and number of vessels by month for all vessels. Vertical dashed lines indicates dates when satellites were added, solid line indicates PIPA closurePIPA closure."}
effort_by_vessel %>% 
  filter(baci_strict) %>%
  group_by(year, gear, treated, post, month, date) %>% 
  summarize(total_hours = sum(hours),
            mean_hours = mean(hours),
            n_mmsi = n()) %>% 
  ungroup() %>% 
  select(-year) %>% 
  gather(variable, value, -c(date, month, post, gear, treated)) %>% 
  ggplot(aes(x = date, y = value, color = treated)) +
  geom_point(shape = ".") +
  facet_wrap(~variable, scales = "free", ncol = 1) +
  ggtheme_plot() +
  scale_color_brewer(palette = "Set1") +
  geom_vline(xintercept = lubridate::date("2015/01/01")) +
  geom_vline(xintercept = lubridate::date("2014/06/01"), linetype = "dashed") +
  geom_vline(xintercept = lubridate::date("2015/12/31"), linetype = "dashed")
```

## Analyses

<!-- antes de 06/01/14 habian 3 sats, despues 6 hasta 12/31/15, luego 10 hasta 2017 y ahora hay muchos jaja -->

The first analysis focuses on identifying the response of fishing vessels to PIPA closure. A spatial closure might cause fishers to modify their behavior as they adapt to a new state of the world. Some may have to travel further distances to find new fishing grounds, which would lead them to fish harder to compensate for the associated extra fuel costs. If fishers had developed experience for fishing in particular sites, being excluded might impose a learning cost on them, as they identify new fishing grounds. This would result in increased search times. However, if the area of influence of a vessel is significantly larger than that of the spatial closure, they might already know other places and times where to fish. Our variables of interests intend to caputre these possible reactions.

We use daily fishing hours, daily proportion of fishing vs. non-fishing hours, and daily distance traveled (km). We compare our variable of interest before and after the implementation of PIPA using a Difference-in-Differences approach, where we track the variable of interest for vessels described in the previous section. Our specification is the following:

$$
y_{i,t} = \alpha + \beta_1 Post_t + \beta_2 Treat_i + \beta_3 Post_t \times Treat_i + \phi_t + \gamma_i + \epsilon_{i,t}
$$

Where $y_{i,t}$ is the variable of interest for vessel $i$ in time period $t$. A dummy variable $Post_t$ takes the value of 0 for all dates prior to PIPA implementation and a value of 1 for all dates including and following PIPA implementation. $Treat_i$ is a dummy variable indicating whether a vessel belongs to the control ($Treat_i = 0$) or treatment ($Treat_i = 1$) group. $\alpha$ is the standard intercept, $\beta_1$ captures the temporal trend change, $\beta_2$ captures the difference between treated and control groups, and $\beta_3$ is our parameter of interest: de DiD estimate capturing the treatment effect. Finally, $\phi_t$ and $\gamma_i$ represent month-, and flag-level dummies that account for seasonality or country-level management interventions\footnote{We test for other more complex specifications that interact a quarterly dummy or year-month dummies with the treatment group and find qualitatively the same results.}.

```{r, model_fitting1}
model_base <- effort_by_vessel %>% 
  filter(baci_strict,
         gear == "purse_seines") %>% 
  mutate(post = post * 1,
         treated = treated * 1,
         year2 = year * year,
         year_month_c = as.character(year_month),
         quarter_c = as.character(quarter),
         flag_fe = !flag %in% c("JPN", "MEX", "NIC", "NZL", "PHL", "USA", "KIR", "PNG"),
         flag_fe_oth = flag_fe & !flag == "OTH",
         sate = case_when(date < lubridate::date("2014/06/01") ~ "1",
                          date > lubridate::date("2015/12/31") ~ "3",
                          T ~ "2"))

# See file combine_vessel_info_and_effort to see how experiment columns are created ~line 70
## Controls are PNA
experiment1 <- model_base %>% 
  filter(experiment1)

## Controls are everything BUT Chinese
experiment2 <- model_base %>% 
  filter(experiment3)

## Controls are Taiwanese
experiment3 <- effort_by_vessel_JPN %>% 
  select(-PNA) %>% 
  mutate(post = post * 1,
         treated = treated * 1,
         year2 = year * year,
         year_month_c = as.character(year_month),
         quarter_c = as.character(quarter),
         flag_fe = F,
         flag_fe_oth = F,
         sate = case_when(date < lubridate::date("2014/06/01") ~ "1",
                          date > lubridate::date("2015/12/31") ~ "3",
                          T ~ "2"),
         baci_strict = T,
         experiment1 = F,
         experiment2 = F,
         experiment3 = F,
         flag = "JPN",
         baci_relaxed = T) %>% 
  rbind(filter(model_base, treated == 1))

## Fit the models

source(here::here("scripts", "did.R"))
source(here::here("scripts", "did_quarter.R"))
source(here::here("scripts", "did_yearmonth.R"))
source(here::here("scripts", "termplot.R"))


## Simple dids for each data
models_exp0_did <- model_base %>% did()
models_exp1_did <- experiment1 %>% did()
models_exp2_did <- experiment2 %>% did()
models_exp3_did <- experiment3 %>% did(flag_fe = F)

# Quarterly did
models_exp0_did_q <- model_base %>% did_quarter()
models_exp1_did_q <- experiment1 %>% did_quarter()
models_exp2_did_q <- experiment2 %>% did_quarter()
models_exp3_did_q <- experiment3 %>% did_quarter()

# Monthly did
models_exp0_did_ym <- model_base %>% did_yearmonth()
models_exp1_did_ym <- experiment1 %>% did_yearmonth()
models_exp2_did_ym <- experiment2 %>% did_yearmonth()
models_exp3_did_ym <- experiment3 %>% did_yearmonth()
```

```{r, eval = T}
regions_few_points <- c("HS FSM 1", "HS MHL 1", "HS MUS 1", "HS NRU 1", "HS PNG 2", "HS SLB 1", "HS TUV 1")

displacement_data <- readRDS(file = here::here("data", "displacement_data_ps.rds")) %>% 
  filter(year < 2018,
         !country %in% regions_few_points)
```

Our second part of the analyses focuses on the redistribution of fishing effort. The role of institutions may play an important role. As stated before, LSMPAs often span the entirety of an EEZ. In the case of purse seining in the PNA, vessels purchase access to the fishery at the beginning of the season. If a vessel decides to fish within the PNA, they've already made the decission to pay for access, expecting higher returns from fishing there than in the High Seas. In the particular case of PIPA, one would expect that a vessel holding a permit to fish in I-Kiribati waters would then move to other fishing grounds within Kiribati after the closure. Alternatively, if a vessel fishes illegaly in I-Kiribati waters before the implementation, one would expect them to reallocate to other regions with equal or less enforcement. In this case, they might then choose to move to other I-Kiribati waters, waters of countries that have similar enforcement levels, or the High Seas.


We discretize spatial units by using a polygon for PIPA \footnote{we would expect to see a decrease here} and distinct spatial units for each EEZ of each country. Some vessels might shift from EEZs into the high seas, but we are interested in knowing *where* in the high seas, so we incorporate additional regions by using a 1 degree buffer of the high seas arround each of the EEZ regions. The rest of the high seas are merged into a single spatial unit. For example, if we were to do this only for Kiribati, we would have 8 spatial units: PIPA, three EEZs, three 1-degree buffers of high seas arround each EEZs, and the rest of the high seas. Whenever the buffers overlapped between themselves, we randocmply clipped one on to the other. EEZs that had sporadic fishing events were pooled into a group of "others".

To evaluate this change in effort allocation, we regress our variable of interest (*i.e.* fishing hours) on the interaction between a dummy variable indicating the policy intervention and a dummy variable for countries. This gives us the by-country change in proportional allocation of fishing effort:

$$
y_{i,t} = \alpha + \beta_1Post + \beta_{2,i}Country + \beta_{3,i}Post_t \times Country_i + \epsilon_{i,t}
$$

Our variable of interest, $y_{i,t}$ represents the proportion of fishing hours that country $i$ receives at time $t$. $Post$ also represents a policy dummy that takes the value of 0 for all dates before implementation of PIPA, and 1 otherwise. $Country$ is a dummy variable for countries for the spatial units defined above. Our parameter of interest is $\beta_{1,i}$, which captures the country-level change in proportional fishing effort.

<!-- Like before, we include a dummy variable for months and allow for a second degree polynomial trend through time. -->

All regression coefficients were estimated via ordinary least squares, and heteroskedastic-robust standard errors were calculated. All analyses were performed in R version `r paste(version$major, version$minor, sep = ".")` \citep{rcore_2018}. Raw data and code used in this work are available on [github](https://github.com/jcvdav/MPA_displacement).

<!-- Figure \ref{fig:redist_trend_spa} shows the yearly average of the relative spatial allocation of fishing efforts; this is the spatial representation of \ref{fig:redist_trend} averaged by year. -->

<!-- Explicar cuales EEZs habia en total y cuales quedaron en un pie de nota o algo asi -->

```{r}
ba_disp <- displacement_data %>% 
  filter(gear == "purse_seines") %>% 
  group_by(country, post) %>% 
  summarize(h_prop_m = mean(h_prop)) %>% 
  spread(post, h_prop_m) %>% 
  mutate(change = (`TRUE` - `FALSE`) * 100) %>% 
  select(country, change)

knitr::kable(x = ba_disp,
             digits = 2,
             booktabs = T,
             caption = "\\label{tab:ba_disp}Changes in the relative allocation of fishing effort by region (EEZ, PIPA, high seas) and gear.") %>%
  kableExtra::kable_styling(latex_options = "HOLD_position")
```

<!-- **An additional test to add to the wish list is to see if vessels are fishing closer to each other following the PIPA closure.** -->

\clearpage

# Results

Our DiD analysis shows an overall increase in purse seine fishing hours, even after accounting for the introduction of new satellites (Table \ref{tab:did}). This coefficient estimate is consistent for different model specifications and across groups of treatment and controls. These effectively represent the patterns observed in Figure \ref{fig:all_vessels}. The $\beta_3$ coefficient indicating our treatment effect suggests that, relative to the control, treated vessels fish less, in the order of 0.5 hours per day. Another way to interpret this is that the increase in fishing effort by treated vessels has ocurred at a lower rate than control vessels. This result is robust and significant ($p < 0.05$) for all model specifications using our main control group, with the effect varying between $\beta_3 = -0.515$ and $\beta_3 = -0.934$. These values are equivalent to a 15 - 28 reduction in fishing hours per month.

Regressions coefficients for our DiD analysis are shown in Table \ref{tab:did}. Columns 1 -3 use all data for controls. Columns 4 - 6 use only PNA vessels as controls, and columns 7 - 9 exclude Chinese vessels. Columns 10 and 11 use Japanese vessels in the Pacific as controls, and hence don't include flag FEs to avoid perfect colinearity between Japanese flag and control\footnote{Results of the same analysis is shown for longliners in \ref{tab:long}}. It must be noted, however, that when reducing the linear structure of the $Pre \times Post$ design and instead interacting the treatment dummy with quarterly or year-month combinations we are not able to reject the null hypotheses of no change (Figures \ref{fig:q1,fig:q2,fig:q3,fig:q4,fig:ym1,fig:ym2,fig:ym3,fig:ym4}).

```{r, echo = F, fig.cap = "\\label{fig:all_vessels}Daily fishing hours for all vessels in our main treatment-control groups. Solid straight lines show a linear trend by period (pre-psot) and treatment. The other red and blue lines show monthly avarages. Vertical dashed line indicates PIPA closure."}
effort_by_vessel %>% 
  filter(baci_strict) %>% 
  mutate(date = lubridate::date(paste(year, month, 15, sep = "/"))) %>%
  ggplot(aes(x = date, y = hours, color = treated)) +
  geom_jitter(shape = ".", height = 0, alpha = 0.1) +
  stat_summary(geom = "line", fun.y = mean, size = 1) +
  geom_smooth(method = "lm", aes(group = paste(treated, post))) +
  ggtheme_plot() +
  scale_color_brewer(palette = "Set1") +
  geom_vline(xintercept = lubridate::date(paste(2015, 01, 01, sep = "/")), linetype = "dashed")
```

\clearpage
\begin{landscape}

```{r, results = "asis"}

models <- c(models_exp0_did,
            models_exp1_did,
            models_exp2_did,
            models_exp3_did)

stargazer::stargazer(models,
                     se = commarobust::makerobustseslist(models),
                     t.auto = T,
                     p.auto = T,
                     intercept.bottom = F,
                     column.sep.width = "1pt",
                     font.size = "footnotesize",
                     type = "latex",
                     omit = c("flag", "month"),
                     add.lines = list(
                       c("Control", rep("All", 3), rep("PNA", 3), rep("-CHN", 3), rep("JPN", 2)),
                       c("Month FE", rep(c("No", "Yes", "Yes"), 4)),
                       c("Flag FE", rep(c("No", "No", "Yes"), 4))),
                     omit.stat = c("adj.rsq", "f", "ser"),
                     header = F,
                     # float.env = "sidewaystable",
                     title = "\\label{tab:did}Difference-in-differences estimates for 3 different controls and 3 different specifications. The first three columns use all data for controls. Columns 4 - 6 use only PNA vessels as controls, and columns 7 - 9 exclude Chinese vessels. Columns 10 and 11 use Japanese vessels in the Pacific as controls, and hence don't include flag FEs. Numbers in parentheses are heteroskedastic-robust standard errors.",
                     out = "DID.html")
```

\end{landscape}
\clearpage

```{r, fig.width = 5, fig.height = 3, fig.cap = "\\label{fig:long}Coefficient estimates for each model. Top pannel indicates variable, x-axis represents model specification, and y-axis coefficient estimate."}
map_df(models,
       commarobust::commarobust_tidy, .id = "model") %>%
  filter(term %in% c("(Intercept)", "post", "treated", "post:treated")) %>%
  mutate(model = as.numeric(model),
         control = case_when(model < 4 ~ "All",
                             model > 3 & model < 7 ~ "PNA",
                             model > 6 & model < 10 ~ "-CHN",
                             T ~ "JPN")) %>% 
  ggplot(aes(x = model, y = est)) +
  geom_errorbar(aes(ymin = est - se, ymax = est + se),
                size = 1,
                width = 0,
                color = "black") +
  geom_point(aes(color = control), size = 3, alpha = 0.7) +
  facet_wrap(~term, scale = "free") +
  ggtheme_plot() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_color_brewer(palette = "Set1")
```

<!-- ## Effort displacement and redistribution -->

Recall that to evaluate the redistribution of fishing effort we only track fishing vessels that belong to the treated group. In this case, we calculate the proportion of fishing effort allocated every month to each spatially explicit region outlined by EEZs and the high seas. For purse seiners, these represent 9 main EEZs, PIPA, the high seas, and a group of other EEZs. Figure \ref{fig:redist_trend_ps} shows the monthly relative fishing hours that each region received by all `r pipa_vessels` treated vessels. The top-left panel shows the change in fishing effort inside PIPA, including the preemptive fishing and immediate reduction previously reported \citep{mcdermott_2018}.

The change in the relative allocation of fishing effort by purse seiners increases in eight of the 12 regions after PIPA implementation (Table \ref{tab:disp_mod}). The largest increase is observed for the I-Kiribati EEZ, with an average increase of 0.11 (*p < 0.001*). In other words, the redistribution of treated vessels caused a 10\% increase in the *relative* allocation of fishing effort within I-Kiribati waters. The only decrease is observed for Papua New Guinea, but the coefficient is not significant. Figure \ref{fig:map_change_ps} provides a spatial representation of these changes. It is evident that the increase in relative fishing effort is greater for for regions closer to PIPA.

```{r, fig.width = 6, fig.height = 9}
mpas_WCPO <- mpas %>% 
  filter(wdpaid %in% c(309888, 400011))

fishing_raster <- readRDS(here::here("data", "fishing_raster_BACI.rds"))

ggplot() +
  geom_raster(data = fishing_raster, mapping = aes(x = lon, y = lat, fill = hours_norm)) +
  geom_sf(data = eez, fill = "transparent", color = "black") +
  geom_sf(data = mpas_WCPO, fill = "transparent", color = "red") +
  # geom_sf(data = small_coast, color = "black") +
  # scale_fill_gradientn(colours = colorRamps::matlab.like2(50), trans = "log10") +
  scale_fill_viridis_c(trans = "log10") +
  facet_grid(year~treated, switch = "y") +
  # coord_equal() +
  ggtheme_map() +
  theme(legend.position = "bottom")
```


```{r, fig.cap = "\\label{fig:redist_trend_ps}Monthly relative allocation of fishing effort by PIPA-fishing vessels before and after PIPA for 9 EEZs, PIPA, the high seas and 'other EEZs'.", eval = T}
displacement_data %>% 
  filter(gear == "purse_seines",
         h_prop > 0) %>%
  ggplot(aes(x = date, y = h_prop, group = post, color = post)) +
  geom_point() +
  geom_vline(xintercept = lubridate::date(paste(2015, 1, 1, sep = "/")), linetype = "dashed") +
  scale_fill_brewer(palette = "Paired") +
  ggtheme_plot() +
  facet_wrap(~country, scales = "free_y") +
  scale_color_brewer(palette = "Set1")
```


```{r, fig.cap = "\\label{fig:map_change_ps}Spatial representation of the mean change in the monthly allocation of fishing effort for purse seiners."}

HS_disp_poly <- matrix(c(200, 10, 210, 10 ,210 ,20 ,200 ,20 ,200, 10),
                  ncol=2,
                  byrow=TRUE) %>%
  list() %>%
  st_polygon() %>% 
  st_sfc(crs = "+proj=longlat +datum=WGS84 +no_defs")

HS_disp <- data.frame(country = "HS",
                      PNA = F) %>% 
  st_set_geometry(HS_disp_poly) %>% 
  mutate(id = country, source = country, PNA = F) %>% 
  select(id, source, PNA, country)

eezs_exclude <- c("HS MUS 1", "EEZ MUS 1", "EEZ MDG 1", "EEZ MDG 2", "HS MDG 1", "HS MDG 2", "HS MOZ 1", "EEZ MOZ 1")

regions <- read_sf(dsn = here::here("data", "spatial", "regions"),
               layer = "regions")

# I AM EXCLUDING MUS WHICH SHOULD BE PUT INTO EEZ
regions %>% 
  rbind(HS_disp) %>% 
  left_join(ba_disp, by = c("id" = "country")) %>%
  filter(!id %in% eezs_exclude) %>%
  ggplot() +
  geom_sf(aes(fill = change), color = "black") +
  ggtheme_plot() +
  scale_fill_gradientn(colors = colorRamps::matlab.like(10))
```

```{r, results = "asis"}
disp_mod_ps_data <- displacement_data %>% 
  filter(!is.na(country),
         gear == "purse_seines",
         !country %in% regions_few_points) %>% 
  mutate(post = post *1,
         year2 = year^2,
         month_c = as.character(month),
         country = fct_relevel(country, "PIPA PIPA 1"),
         sate = case_when(date < lubridate::date("2014/06/01") ~ "1",
                          date > lubridate::date("2015/12/31") ~ "3",
                          T ~ "2"))

disp_mod_ps1 <- lm(h_prop ~ post*country + sate, data = disp_mod_ps_data)


commarobust::commarobust_tidy(disp_mod_ps1) %>% 
  mutate(b = str_detect(term, "Intercept"),
         c = str_detect(term, "post"),
         d = str_detect(term, "sate"),
         a = b*c,
         p1 = ifelse(p < 0.1, "*", ""),
         p2 = ifelse(p < 0.05, "*", ""),
         p3 = ifelse(p < 0.01, "*", ""),
    h_prop = paste0(formatC(est, digits = 3, format = "f"),
                    " (",
                    formatC(se, digits = 3, format = "f"),
                    ")",
                    p1,
                    p2,
                    p3)) %>% 
  filter(b+c+d > 0) %>% 
  select(term, h_prop) %>% 
  knitr::kable(booktabs = T,
               caption = "Change in the relative allocation of purse seining fishing hours by region (R2 = 0.36, F(47, 69000) = 961.1, p < 0.001)") %>% 
  kableExtra::kable_styling(latex_options = "HOLD_position")

```

\clearpage

# Discussion

Our findings provide interesting insights into the effect that LSMPAs can have on vessel behavior and the redistribution of fishing effort. These collection of results shows that the implementation of PIPA caused treated vessels to reduce their fishing hours, and that this effect is greater for purse seiners than longliners. Even though treated vessels fish less, their relative allocation of fishing hours increased for all other fishing grounds. This does not imply that there is more fishing effort exerted by treated vessels, but rather that each region receives a greater portion of the post-PIPA fishing effort of these same vessels, which is lower than pre-PIPA levels. In this section we discuss the implications of vessel-level reductions in fishing effort and the increase in relative allocation of the remaining effort through space. We also provide plausible explanations as to why purse seiners seem to be more reactive to the spatial closure.

A major shortcoming of our analyses is that we do not observe catches or revenues, which ultimately are the factors that guide the decision making process of profit-maximizing agents. Therefore, it is difficult to know whether the reduction in fishing effort represents a positive or negative impact. A decrease in fishing effort is associated to an increase in catches (and therefore greater CPUE) only when the entire fleet does it, and if previous levels of effort were greater than $F_{MEY}$ (*i.e.* the effort that would yield the maximum economic yield). Therefore, it is plausible that the reduction of fishing hours is not done by choice, but rather results from fishers having to increase search time. Upon being relocated, fishers may not identify the best fishing grounds as easily as before, and therefore invest a greater proportion of their time searching for their catch. Further analysis of temporal trends in non-fishing hours, as well as distance traveled should provide us with insights as to why fishers reduced fishing hours.

<!-- **We should have a discussion (over the phone and in the paper) about whether a decrease in fishing hours represents a positive or negative impact of PIPA. Without data on catch it might be hard to really know.** -->

Previous studies on insular environments suggest that vessels move to distant places, which might be translated as increased costs \citep{stevenson_2013}. Nevertheless, they do not use counterfactuals that could help account for system- or fleet-level changes that occur through time. Others have used similar satellite-tracking systems to show that fishing effort accumulates near the edges of spatial closures, yielding greater catches \citep{murawski_2005}. Yet, these vessel tracks do not cover the pre-reserve period, making it difficult identify the contribution of spatial closures to the observed spatial distribution of fishing vessels. Recent work by \citet{elahi_2018} identified that total fishing effort in a focal region where a short-term MPA was implemented showed little change, likely indicating that fishers redistributed fishing effort to compensate for the reduction in available space. Our data is assembled in a similar way, with fishing positions before and after the implementation of PIPA and vessels grouped into treated and control groups. Our BACI design, along with our difference-in-differences analysis allows us to make causal inferences about the effect that large scale marine protected areas have on fishing effort.

The different responses observed between purse seiners and longliners might have two possible explanations. It is likely that PIPA did not contain habitat that longliners would consider optimal. Therefore, the sporadic fishing events that occurred there are of little importance to the fleet, and it is unlikely that the implementation of PIPA has an effect on them. Alternatively, the differences may be due to the nature of each fishing gear. Purse seiners are often constrained by seafloor and termocline depth, and have a smaller spatial footprint \citep{kroodsma_2018}. Tuna purse seiners are known to have greater proportion of null sets (*i.e.* where purse seines effectively cast their nets, but no catch is obtained) during El Niño years, where the termocline deepens in the Eastern Pacific \citep{dreyfusleon_2015}. On the other hand, longliners may be more flexible as to where they can deploy their longlines. \citet{ortuocrespo_2018} evaluated the ecological niche of the pelagic longline fleet, and suggest that the fleet may be under-utilizing the ocean, meaning that they can easily redistribute elsewhere.

Our work suggests that the implementation of LSMPAs can have important implications for purse seiners, and less so for longliners. We also show that fishing effort is redistributed to areas close by. Future management interventions that aim to close large portion of the oceans should consider how fishing effort will change in space and through time, and the ecological implications of this redistribution to ensure that fishing effort is not just displaced elsewhere, leading to overfishing in adjacent waters.

Furthermore, a growing body of literature suggests that closing the high seas to all fishing could increase fishery yields and profitability of fisheries, with negligible costs to food security \citep{white_2014,sumaila_2015,sala_2018a,schiller_2018}.

<!-- # Notes and things to do -->

<!-- Additionally, run similar regressions on distance traveled, and total vessel hours (not just fishing hours), that could account for the extended "search" time. -->

\clearpage

<div id = "refs"></div>

\clearpage

# Appendix

\beginsupplement

## Stream of data by vessel, group, and period

```{r, fig.width = 8, fig.height = 6, fig.cap = "\\label{fig:baci_strict}Stream of fishing events by vessels through time. Each line represents a vessel, with dots indicating months with fishing activity and colors indicating the pre and post periods.", eval = T}
filter(effort_by_vessel, baci_strict) %>% 
  mutate(gear = ifelse(gear == "purse_seines", "Purse seiners", "Drifting longliners")) %>% 
  group_by(mmsi, date, gear, post, treated) %>%
  summarize(h = sum(hours, na.rm = T)) %>%
  ungroup() %>% 
  arrange(treated, post, mmsi) %>%
  mutate(vessel = group_indices(., paste(gear, treated, mmsi)),
         treated = ifelse(treated, "Treated", "Control")) %>% 
  ggplot(aes(
    x = date,
    y = vessel,
    fill = post,
    color = post,
    group = paste(vessel, post)
    )) +
  geom_point(size = 1) +
  theme_bw() +
  facet_wrap(~ treated, scales = "free_y") +
  ggtheme_plot() +
  scale_color_brewer(palette = "Set1") +
  scale_y_continuous(breaks = NULL, labels = NULL) +
  theme(legend.position = "none")
```

\clearpage

## Longliners

Longliners show a similar pattern of effort reduction. However, the magnitude of the $\beta_3$ coefficient is smaller (ranging from  $\beta_3 = -0.98$ to $\beta_3 = -1.125$) and not significant unless including month and flag FEs for the general treatment-control groups, or when excluding Chinese vessels. This, along with higher standard error values suggest that longliners have a smaller and more variable response to the implementation of LSMPAs.

```{r, results = "asis"}
model_base <- effort_by_vessel <- readRDS(file = here::here("data", "effort_by_vessel.rds")) %>% 
  filter(year < 2018,
         !gear == "purse_seines",
         baci_strict) %>% 
  mutate(post = post * 1,
         treated = treated * 1,
         year2 = year * year,
         year_month_c = as.character(year_month),
         quarter_c = as.character(quarter),
         flag_fe = !flag %in% c("JPN", "MEX", "NIC", "NZL", "PHL", "USA", "KIR", "PNG"),
         flag_fe_oth = flag_fe & !flag == "OTH",
         sate = case_when(date < lubridate::date("2014/06/01") ~ "1",
                          date > lubridate::date("2015/12/31") ~ "3",
                          T ~ "2"))

# See file combine_vessel_info_and_effort to see how experiment columns are created ~line 70

## Controls are everything BUT Chinese
experiment2 <- model_base %>% 
  filter(experiment3)


## Simple dids for each data
ll_models_exp0_did <- model_base %>% did()
ll_models_exp2_did <- experiment2 %>% did()

models <- c(ll_models_exp0_did,
            ll_models_exp2_did)

stargazer::stargazer(models,
                     se = commarobust::makerobustseslist(models),
                     t.auto = T,
                     p.auto = T,
                     intercept.bottom = F,
                     column.sep.width = "1pt",
                     font.size = "small",
                     type = "latex",
                     omit = c("flag", "month"),
                     add.lines = list(
                       c("Control", rep("All", 3), rep("-CHN", 3), rep("JPN", 2)),
                       c("Month FE", rep(c("No", "Yes", "Yes"), 3)),
                       c("Flag FE", rep(c("No", "No", "Yes"), 3))),
                     omit.stat = c("adj.rsq", "f", "ser"),
                     header = F,
                     # float.env = "sidewaystable",
                     title = "\\label{tab:long}Difference-in-differences estimates for 2 different controls and 3 different specifications. The first three columns use all data for controls. Columns 4 - 6 exclude Chinese vessels. Numbers in parentheses are heteroskedastic-robust standard errors.",
                     out = "DID_ll.html")
```

\clearpage

## Other model specifications for purse seiners

We include a second degree polinomial for years as:

$$
y_{i,t} = \alpha + \beta_1 Post_t + \beta_2 Treat_i + \beta_3 Post_t \times Treat_i + \phi_t + \gamma_i + \omega Y_t + \Omega Y_t^2 + \epsilon_{i,t}
$$

```{r, results = "asis"}
model_data <- readRDS(file = here::here("data", "effort_by_vessel.rds")) %>%
  filter(year < 2018,
         baci_strict,
         gear == "purse_seines") %>% 
  mutate(post = post * 1,
         treated = treated * 1,
         mmsi_c = as.factor(mmsi))

m1 <- lm(hours ~ post*treated, data = model_data)
m2 <- lm(hours ~ post*treated + month_c, data = model_data)
m3 <- lm(hours ~ post*treated + month_c + year_c, data = model_data)
m4 <- lm(hours ~ post*treated + month_c + year_c + flag, data = model_data)

models <- list(m1, m2, m3, m4)

stargazer::stargazer(models,
                     se = commarobust::makerobustseslist(models),
                     t.auto = T,
                     p.auto = T,
                     type = "latex",
                     omit = c("flag", "year", "month"),
                     add.lines = list(c("Month FE", "No", "Yes", "Yes", "Yes"),
                                      c("Year FE", "No", "No", "Yes", "Yes"),
                                      c("Flag FE", "No", "No", "No", "Yes")),
                     omit.stat = c("adj.rsq", "f", "ser"),
                     header = F,
                     title = "\\label{tab:purse}Fishing hours from GFW for purse_seiners. Asterisks indicate significance levels. Numbers in parentheses represent heteroskedastic-robust standard errors.")
```

### Quarterly DID interactions

```{r, fig.cap = "\\label{fig:q1}Interaction of quarters and treatment. Control group is all vessels."}
# Quarterly did
termplot(models_exp0_did_q, q = T)
```

```{r, fig.cap = "\\label{fig:q2}Interaction of quarters and treatment. Control group is vessels from PNA."}
termplot(models_exp1_did_q, q = T)
```

```{r, fig.cap = "\\label{fig:q3}Interaction of quarters and treatment. Control group excludes Chinese vessels."}
termplot(models_exp2_did_q, q = T)
```

```{r, fig.cap = "\\label{fig:q4}Interaction of quarters and treatment. Control group is japanese vessels."}
termplot(models_exp3_did_q, q = T)
```

### Year-month DID interactions

```{r, fig.cap = "\\label{fig:ym1}Interaction of year-month and treatment. Control group is all vessels."}
# Monthly did
termplot(models_exp0_did_ym)
```

```{r, fig.cap = "\\label{fig:ym2}Interaction of year-month and treatment. Control group is vessels from PNA."}
termplot(models_exp1_did_ym)
```

```{r, fig.cap = "\\label{fig:ym3}Interaction of year-month and treatment. Control group excludes Chinese vessels."}
termplot(models_exp2_did_ym)
```

```{r, fig.cap = "\\label{fig:ym4}Interaction of year-month and treatment. Control group is japanese vessels."}
termplot(models_exp3_did_ym)
```




\clearpage
